---
title: "Using Regular Expressions in R"
author: "ZevRoss Spatial Analysis"
date: "April 21, 2016"
output: html_document
---

## Regular Expressions
We use regular expressions frequently enough in our line of work but they're one of those things that I constantly need to refer to tutorials or references to ...


## Special Characters
This table was copied from:
https://bioinfomagician.wordpress.com/2013/10/27/regular-expression-tutorial-1-special-characters/

Symbol			|	Name			|	Function
------------- 	|	------------- 	|	------------- 
\\    			|	back slash		|	escape character
[ ]  			|	square brackets |  	single character match
{ }  			|	curly braces	|  	repeats
( )  			|	parenthesis 	|   reference or subexpression
^    			|	hat             |	beginning of a line (not string)
$    			|	dollar          |	end of a line (not string)
\|    			|	pipe            |	alternation [OR]
*    			|	asterisk        |  	zero or more times of repeat
+    			|	plus sign       |  	1 or more times of repeat
?    			|	question mark   |	occur 0 times or once 
.    			|	dot             |  	any single character
!    			|	exclamation     |	negation [NOT]



Symbol			|	Name			
------------- 	|	------------- 	
\\t				|	tab
\\n				|	new line
\\r				|	carriage return
\\f				|	form feed character (end of page character)


### Identify certain character class

Symbol			|	Name			
------------- 	|	-------------
\\s				|	a white space
\\S				|	non-white space
\\d				|	a digit [0-9]
\\D				|	non-digit [^0-9]
\\w				|	word character [a-z, A-Z]
\\W				|	non-word character
\\A				|	beggining of a string
\\z				|	end of a string









```{r}
x<-c("145_bus23", "230_busy731", "AFQ_247_bustedje", "ll4bus", "Bus63_x391")

grep(pattern = "bus", x = x)

grep(pattern = "bus$", x = x)

grep(pattern = "bus", x = x, ignore.case = T)

```










You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
